---
title: "Homework 4"
author: "Aolong Zhang"
date: "2023-10-06"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r}

library(readxl)
Tesla_sales <- read_excel("C:/Users/17215/OneDrive/Desktop/Tesla sales.xlsx")
View(Tesla_sales)

class(Tesla_sales)

## convert dataset to time series
data1 = ts(Tesla_sales$`Total Sales`, start = c(2015),end = c(2023,6),
           frequency = 12)

library(fpp)
library(fpp2)
library(TTR)

## dataset infomration
attributes(Tesla_sales)
plot(data1,ylab= "Total Sales")

## acf table shows there is strong increasing trend
acf(data1)


## take mean of all avaible history
mean_forecast = meanf(data1,5)
plot(mean_forecast)



## naive
naive_forecast = naive(data1,5)
plot(naive_forecast)
naive_forecast

##random walk
rwf_forecasting = rwf(data1,5)
rwf_forecasting = rwf(data1,5,drift = TRUE)
plot(rwf_forecasting)


## Seasonal naive
snaive_forecast = snaive(data1,5)
plot(snaive_forecast)


## Moving average (5 and 9)
MA5_forecast = ma(data1,order = 5)
MA9_forecast = ma(data1,order = 9)
MA5 = forecast(MA5_forecast,h=5)
MA9 = forecast(MA9_forecast,h=5)


##plot all line in the chart
plot(mean_forecast)
lines(naive_forecast$mean,col="sienna")
lines(rwf_forecasting$mean,col="green")
lines(snaive_forecast$mean,col="orange")
lines(MA5_forecast,col="purple")
lines(MA9_forecast,col="red")



## decomposition

ets_forecast = ets(data1)
plot(ets_forecast)

ets_forecast$mse


##SSE
SSE_Simple = HoltWinters(data1,beta = FALSE,gamma = FALSE)
plot(SSE_Simple)
SSE_Simple$SSE

## SSE with trend
SSE_trend = HoltWinters(data1,gamma = FALSE)
plot(SSE_trend)
SSE_trend$SSE


## Holts winters
SSE_winters = HoltWinters(data1)
plot(SSE_winters)
SSE_winters$SSE

##Forecast
forecast_ets = forecast(ets_forecast,h=5)
plot(forecast_ets)

forecast_HW = forecast(SSE_winters,h=5)
plot(forecast_HW)

accuracy(snaive_forecast)
accuracy(mean_forecast)
accuracy(rwf_forecasting)
accuracy(naive_forecast)
accuracy(MA5)
accuracy(MA9)
accuracy(forecast_ets)
accuracy(forecast_HW)


```

#Question 1  
Did 7 differnt models: naive, mean, seasonality,random walk, moving averge 5 period and 9 period,exponential smoothing,and Holts winters.

#Question 2  
all my models are to predict next 5 month sales.For naive,mean,moving average, and random walk, those model did nto include any trend and sensonality.For the seasonality model, becasuse from the acl chart we see the data has no seasonal pattern,the forecsating is not very accuret, you can also see from the chart,the organge line.For ets model,I got ETS(A,N,N),which means the data is simple additive model without any trend or seasonality.For Holts Winters,I tried 3 situations 1.without beta and gamma, 2.with only beta and alpha,and 3.defult.And turns out the defult is better with smallest SSE.



#question 3  

```{r}
plot(mean_forecast)
lines(naive_forecast$mean,col="sienna")
lines(rwf_forecasting$mean,col="green")
lines(snaive_forecast$mean,col="orange")
lines(MA5_forecast,col="purple")
lines(MA5$mean,col="purple")
lines(MA9$mean,col="red")
lines(MA9_forecast,col="red")
lines(forecast_HW$mean,col="grey")
lines(forecast_ets$mean,col="khaki")
```



#Question 4  
I choose MAE(mean absolute error) as the indicator to assess the accuracy for each model, and the result is the moving average(9 period) is the best one with the minimum MAE.
